---
title: "ANLT_Assignment 1"
author: "Saetbyeol Choi and Wenbo Yu"
output: 
  html_document:
    toc: true
    toc_float: true
date: "2022-11-20"
---

# a. State the null and alternative hypotheses.
Null Hypothesis

H~0~: $\mu$~0~ = 16.7

Alternative Hypothesis

H~a~: $\mu$~0~ ≠ 16.7

Note that this is a two-sided hypothesis.

# b. Determine the probability of a Type I error.
Type I error: rejecting the null hypothesis when it is in fact true.

$\alpha$ = 0.05

Thus, the probability of a Type I error is 0.05.

# c. Determine the probability of a Type II error. 
```{r}
(z_left = qnorm(0.05/2, mean=16.7, sd=6.0/sqrt(100), lower.tail=T))
```
```{r}
(z_right = qnorm(0.05/2, mean=16.7, sd=6.0/sqrt(100), lower.tail=F))
```
```{r}
(beta = pnorm(z_right, mean=15.5, sd=6.0/sqrt(100), lower.tail=T) -  pnorm(z_left, mean=15.5, sd=6/sqrt(100), lower.tail=T))
```

# d. Simulate 1,000 samples, each of size 100.
```{r}
# solution 1
set.seed(1)
sample1 = replicate(1000, rnorm(100, 15.5, 6.0), simplify = FALSE)
length(sample1)
head(sample1,1)
```
```{r}
# solution 2
set.seed(1)
Organizer <- list()
for (i in 1:1000){
  sample <- round(rnorm(100, mean = 15.5, sd = 6),1)
  Organizer[[i]] <- sample
}
length(Organizer)
head(Organizer,1)
```

# e. Determine the mean of each sample in part (d).
```{r}
# solution 1
head(sapply(sample1, mean),3)
```
```{r}
# solution 2
Organizer_2 <- list()
for(i in 1:1000){
  xmean <- mean(Organizer[[i]])
  Organizer_2[[i]] <- xmean
}
head(Organizer_2,3)
```

```{r}
# cf).Another way to simulate sample & mean of each sample
n = 100
mean = 15.5
sd = 6.0

sample=list()
samplemean=NULL
sdsample=NULL

for(i in 1:1000){
  sample[[i]] = rnorm(n,mean,sd)
  samplemean[i]=mean(sample[[i]])
}

head(sample,1)
head(samplemean,1)
```

# f. For the 1,000 samples obtained in part (d), about how many would you expect to lead to nonrejection of the null hypothesis? Explain your answer. 
```{r}
beta*1000
```

# g. For the 1,000 samples obtained in part (d), determine the number that lead to nonrejection of the null
hypothesis.
```{r}
# solution 1
sum(samplemean > z_left & samplemean < z_right)
```
```{r}
# solution 2 
reject_number <- 0
for(i in 1:1000){
  if (Organizer_2[[i]] < 15.52402 || Organizer_2[[i]]> 17.87598){
    reject_number <- reject_number + 1
  }
}
1000-reject_number
```

# h. Compare your answers from parts (f) and (g), and comment on any observed difference.
The expected number of non-rejection is 483.9947 and the non_reject number is 483(solution1)/456(solution2).
They are quite close on the slope of 1000.
The difference is quite normal, it's like an uneven coin-tossing problem having a probability of head equal to 0.4839947 and probability of tail equal to (1-0.4839947). If we extend the number of samples, the percentage of non-rejection will be closer to 0.4839947.

# i. Plot the power curve for the range of true µ between 14 to 19. Interpret your plot.
```{r}
mu = 16.7
popSigma = 6.0
n = 100
alpha = 0.05
# Critical point
c = qnorm(alpha, mean=mu, sd=popSigma/sqrt(n))
trueMu = seq(14,19,by=0.1)
beta = pnorm(z_right, mean=trueMu, sd=popSigma/sqrt(n)) - pnorm(z_left, mean=trueMu, sd=popSigma/sqrt(n))
power = 1-beta

plot(trueMu, power, 'l',
     xlab=expression(mu), 
     main=expression(paste("Power Curve (",alpha, "=0.05, n=100)")))
```
